\section{Results}
\subsection{Single Phone Model}
Using the architecture detailed in section 3, we trained the network for 10 epochs
and measured its effectiveness using $10$-fold cross-validation.
Training/validation data was compiled from the diary study conducted with the Nexus 5X.

The average accuracy across all folds was $92\%$, and the corresponding
confusion matrix is shown in Table ~\ref{fig:confusion1}.
 The network is able to learn how to  classify the Table state very accurately, which is understandable given  the consistent nature of the state (i.e. always in a still, flat position). 
The states of Backpack and Pocket were also classified with pretty moderate accuracy,
with the misclassifications likely stemming from the more diverse nature of 
the two states (i.e. the user could have been moving or still when the phone was
in their backpack or pocket). 
The Hand state was classified with the least accuracy,
and misclassified most often instead as the Backpack state. 
These misclassifications likely also stem from the varied positions of the Hand state, 
 as in addition to the user using the phone actively in their hand, 
 they may also walk with their phone in hand
 but just to hold the phone and not actually using it.
 
One problem our classifier had was in distinguishing between cases where the phone was flat on a table or still in a non-moving backpack.
For future work, we propose a post-processing step after obtaining the classification output.
In this step, classifications from timeframes before and after the one in question would be considered to help `smooth' the outputs.

\begin{table}[h]
\caption{Confusion matrix of the network predictions on Diary Study 1 (Nexus 5X). Each entry indicates the percent of
total instances that were predicted as the predicted class by the network and labeled the actual class.}\label{fig:confusion1} \centering
\begin{tabular}{ l rrrr }  
\toprule
      &  \multicolumn{4}{c}{\textit{Actual}} \\
\cmidrule{2-5}
\textit{Predicted}		&	Backpack    & 	Pocket 	& 	Hand	&	Table \\
\midrule
Backpack			&	26.6\% 	&	 .5\%		&	2.4\% 	&	2.9\% \\
Pocket			&	1.6\% 	&	4.1\% 	&	 .1\% 	&	 .1\% \\
Hand			&	1.7\% 	&	 .3\% 	&	3.5\% 	&	 .6\% \\
Table			&	 .4\% 	&	 .2\% 	&	 .2\% 	&	96.5\%\\
\bottomrule
\end{tabular}
\end{table}


\subsection{Multiple Phone Models}
We also attempted to validate our networks across different phone models, by training
on data collected by one phone model and then validating on data collected on a 
different phone model. Specifically, we first applied the calibration process described in section 4.3 
to both the Nexus 5X diary study data (Diary Study 1) and Nexus 5 diary study data (Diary Study 2 and Diary Study 3).
We then trained a network of our proposed architecture on the Nexus 5X data for 20 epochs, 
and validated the network on the two separate Nexus 5 diary studies.

Validating on the first Nexus 5 diary study (Diary Study 2), our network
had an accuracy of $78\%$, significantly lower than the accuracy observed
in the single phone model case. The  confusion matrix shown in table~\ref{fig:confusion2}.

\begin{table}[h]
\caption{Confusion matrix of the network predictions on Diary Study 2 (Nexus 5).}\label{fig:confusion2}
\centering
\begin{tabular}{ l rrrr }  
\toprule
   & \multicolumn{4}{c}{\textit{Actual}} \\
\cmidrule{2-5}
\textit{Predicted}	&	Backpack    & 	Pocket 	& 	Hand	&	Table \\
\midrule
Backpack			&	34.2\% 	&	1.6\%	&	1.7\% 	&	13.1\% \\
Pocket			&	.1\% 		&	18.5\% 	&	.0\%		&	.0\% \\
Hand			&	5.7\%	&	.0\%	 	&	.4\% 		&	.0\% \\
Table			&	.0\% 		&	.3\% 		&	.0\%	 	&	24.5\%\\
\bottomrule
\end{tabular}
\end{table}

Validating the second diary study, with the Nexus 5 (Diary Study 3), our trained network had a validation
accuracy of $91\%$--an accuracy much closer to the cross validation accuracy
demonstrated in the single phone model case. The  confusion matrix is 
shown in table~\ref{fig:confusion3}.

\begin{table}[h]
\caption{Confusion matrix of the network predictions on Diary Study 3 (Nexus 5).}\label{fig:confusion3} \centering
\begin{tabular}{ l rrrr }  
\toprule
 & \multicolumn{3}{c}{\textit{Actual}} \\ 
\cmidrule{2-5}
\textit{Predicted}	&	Backpack    & 	Pocket 	& 	Hand	&	Table \\
\midrule
Backpack			&	5.3\% 	&	4.0\%	&	1.3\% 	&	.1\% \\
Pocket			&	  .6\% 	&	3.7\% 	&	.5\% 		&	.1\% \\
Hand			&	 .0\% 	&	1.4\% 	&	5.5\% 	&	.1\% \\
Table			&	 .9\% 	&	 .3\% 	&	.0\%		&	76.3\%\\
\bottomrule
\end{tabular}
\end{table}


The disparity in accuracy between the two different diary studies with the Nexus 5 was unexpected,
since both were conducted within the same Nexus 5 device. We did not have time to explore the root
cause for this disparity, so we cannot conclude definitely if a single network is applicable across phone models. 
The results from Diary Study 3 ($91\%$ accuracy) suggest that a linear calibration
strategy may be effective, but further investigation is needed for definitive understanding.
